Value-Suppressing Uncertainty Maps

### 2017-03-27 paper/template.tex
Typos

In some cases, this second assumption is violated. For instance, an analyst might be interested in ``long tail risks'' or other ``black swan events,'' where the impact of an value, no matter how uncertain, must be considered and planned for~\cite{taleb2011black}. Other analysis tasks (such as filtering out outliers), require increased, rather than decreased, discrimbinability when uncertainty low. However, for many information fusion tasks, the assumption is that uncertainty is related to data quality, or the uncertainty of particular data values \cite{riveiro2007evaluation}. As with null hypothesis significance testing, the aim would then be to avoid Type I errors, where patterns arising from statistical noise or measurement error are interpreted as significant signals.

@@

Another reason for limiting our discussion of bivariate mappings to color and size channels is that the perceptual distinguishability of these channels has been well-studied by prior work. CIELAB and other color spaces are designed with the intention of perceptual uniformity--- that is, distance in the color space should universally correspond with perceptual distance. Recent work~\cite{stone2014engineering,szafir2014adapting} has investigated the discriminability of colors in such spaces not only under laboratory conditions, but also among the disparate display conditions and color vision proficiency of crowdworkers. Similarly, formative work by Cleveland \& McGill~\cite{cleveland1984graphical} assesses the discriminability of marks of different height, and this too has been extended both to the more variable population of crowdworkers, and to the area channel specifically~\cite{heer2010crowdsourcing,talbot2014four}. We use values from these works as benchmarks for selecting thresholds of discriminability. The ``just noticeably difference'' (JND) of these channels provides a useful minimum; in practice, we would wish to create maps with marks that are far more reliably distinguishable than the 50\% or 75\% discriminability rate of a JND. Figure \ref{fig:performance} shows how these benchmarks can be used to select reasonable thresholds. Other factors, such as nameability and aesthetic preference, can be added in as additional weights impacting the desirability of certain color palettes, as in Gramazio et al.~\cite{gramazio2017colorgorical}.

How quickly the number of bins decreases is another parameter of VSUMs. If this degradation is too large, then we may only have a few discriminable levels of uncertainty. If this degradation is too small, then we can be overwhelmed with colors. For instance, if we only decrement by one bin per level of uncertainty, there would be \emph{1,485} different colors in a VSUM with a tolerance thresold of $1$ unit in CIELAB space. In the examples in this paper,  we halve the number of bins at each increasing level of uncertainty. When the initial bins at the highly certain region is a power of $2$ (as in Fig. \ref{fig:example}), this gives a tree structure to the resulting VSUM. As uncertainty increases, bins in the previous level map to one and only one ``parent'' in the next uncertainty level, resulting in fewer crossings of color boundaries among rows.

>>>>

In some cases, this second assumption is violated. For instance, an analyst might be interested in ``long tail risks'' or other ``black swan events,'' where the impact of an value, no matter how uncertain, must be considered and planned for~\cite{taleb2011black}. Other analysis tasks (such as filtering out outliers), require increased, rather than decreased, discriminability when uncertainty low. However, for many information fusion tasks, the assumption is that uncertainty is related to data quality, or the uncertainty of particular data values \cite{riveiro2007evaluation}. As with null hypothesis significance testing, the aim would then be to avoid Type I errors, where patterns arising from statistical noise or measurement error are interpreted as significant signals.

@@

Another reason for limiting our discussion of bivariate mappings to color and size channels is that the perceptual distinguishability of these channels has been well-studied by prior work. CIELAB and other color spaces are designed with the intention of perceptual uniformity--- that is, distance in the color space should universally correspond with perceptual distance. Recent work~\cite{stone2014engineering,szafir2014adapting} has investigated the discriminability of colors in such spaces not only under laboratory conditions, but also among the disparate display conditions and color vision proficiency of crowdworkers. Similarly, formative work by Cleveland \& McGill~\cite{cleveland1984graphical} assesses the discriminability of marks of different height, and this too has been extended both to the more variable population of crowdworkers, and to the area channel specifically~\cite{heer2010crowdsourcing,talbot2014four}. We use values from these works as benchmarks for selecting thresholds of discriminability. The ``just noticeable difference'' (JND) of these channels provides a useful minimum; in practice, we would wish to create maps with marks that are far more reliably distinguishable than the 50\% or 75\% discriminability rate of a JND. Figure \ref{fig:performance} shows how these benchmarks can be used to select reasonable thresholds. Other factors, such as nameability and aesthetic preference, can be added in as additional weights impacting the desirability of certain color palettes, as in Gramazio et al.~\cite{gramazio2017colorgorical}.

How quickly the number of bins decreases is another parameter of VSUMs. If this degradation is too large, then we may only have a few discriminable levels of uncertainty. If this degradation is too small, then we can be overwhelmed with colors. For instance, if we only decrement by one bin per level of uncertainty, there would be \emph{1,485} different colors in a VSUM with a tolerance threshold of $1$ unit in CIELAB space. In the examples in this paper,  we halve the number of bins at each increasing level of uncertainty. When the initial bins at the highly certain region is a power of $2$ (as in Fig. \ref{fig:example}), this gives a tree structure to the resulting VSUM. As uncertainty increases, bins in the previous level map to one and only one ``parent'' in the next uncertainty level, resulting in fewer crossings of color boundaries among rows.
